Demo arXiv template

### 2022-11-28 paper/paper.qmd
section on distribution fit and tweak in normalising plot

$$F_{\eta}(x(\mathbf{s}; \mathbf{t})),$$ where $\eta \in H \subseteq \mathbb{R}^{d_{\eta}}$ is the distribution parameter and $d_{\eta}$ is the number of parameter of $\eta$.

Considerations need to be given on selecting the best distribution to fit, 
Various discussion on the fitting methods and goodness of fit. 
Potentially, more care about how much difference each combination give in terms of the index value.

sometimes, to account for the monthly variation in the data, the fitting is done separately for each month and the noramlising step can be written as $\Phi^{-1}[F_{\eta}^m(x(\mathbf{s};t_{j^*}))]$, where $j^*$ is all the indexes that satisfy $j^* \mod 12 = m$ for each $m = 0, 1,\cdots, 11$.

- In some cases, the distibution and inverse normal can be multivariate, i.e. use copula.

@@

This step maps the univariate series into a different scale, typically for the ease of comparison. For example, a normal scale, [0, 1], [0, 100] may be favored to report certain indexes. In drought indexes, i.e. SPI or SPEI, the quantiles from the fitted distribution is converted into the normal scale via the normal reverse CDF function: $\Phi^{-1}(.)$.

Normalising is usually used in the end of the pipeline and its main difference from the scaling step is that here the change of scale also changes the distribution of the variable. 
While being commonly used, this step can get criticism from analysts for forcing the data into the decided scale, which can be either unnecessary or inaccurately exaggerate or downplay the outliers. Also, the use of normal scale needs to be interpreted with caution. @fig-normalising illustrates the normal density not being directly proportional to its probability of occurrence. This is concerning, especially at the extreme values, where a normal density of -5 and -4 has magnitudes of difference into its probability of occurrence (1 in 3,000,000 and 1 in 31,000, respectively).

```{r fig-normalising}
#| fig-cap: "scatterplot of quantile vs. its density values."
library(tidyverse)
all <- tibble(x = seq(-5, 5, 0.01), y = pnorm(x)) 
dt <- tibble(x = c(-5, -4), y = pnorm(x)) 
label <- tibble(x = c(-5, -4), y = pnorm(x), 
                label = c("a chance \nof 1 in \n 3,488,556", "a chance \nof 1 in \n 31,574"))

dt %>% 
  ggplot() + 
  geom_point(aes(x= x, y =y)) + 
  geom_line(data = all, aes(x = x, y =y)) + 
  ggrepel::geom_label_repel(
    data = label, aes(x = x, y = y, label =label), 
    nudge_y = 0.2, arrow = arrow(length = unit(0.08, "inches")), min.segment.length = 0) + 
  scale_x_continuous(breaks = seq(-5, 5, 1)) + 
  xlab("Density") + 
  ylab("Quantile") + 
  theme_bw() + 
  theme(panel.grid.minor = element_blank())
```

>>>>

Distribution fit can be seen as model fitting in its simplest term. It can be represented by

<!-- With a probability model imposed,  -->

$$F_{\eta}(x(\mathbf{s}; \mathbf{t})),$$ where $\eta \in H \subseteq \mathbb{R}^{d_{\eta}}$ is the distribution parameter and $d_{\eta}$ is the number of parameter of $\eta$. A distribution fit typically aims at finding the distribution that best fit the data. 
Analysts may start from a pool of candidate distributions with a chosen fitting method and goodness of fit measure. 
While it is useful to find the ultimate best distribution to fits the data, from a probabilistic perspective, the fitting procedure itself has uncertainty associated with the data fed and parameter chosen. A reasonable alternative is to understand how much the index values can vary given different distributions, fitting methods, and goodness of fit tests, and whether these variations are negligible in a given application.

@@

This step maps the univariate series into a different scale, typically for the ease of comparison across regions. For example, a normal scale, [0, 1], or [0, 100] may be favored for reporting certain indexes. In drought indexes, i.e. SPI or SPEI, the quantiles from the fitted distribution is converted into the normal scale via the normal reverse CDF function: $\Phi^{-1}(.)$.

Normalising is usually used in the end of the pipeline and its main difference from the scaling step is that here the change of scale also changes the distribution of the variable. 
While being commonly used, this step can get criticism from analysts for forcing the data into the decided scale, which can be either unnecessary or inaccurately exaggerate or downplay the outliers. Also, the use of normal scale needs to be interpreted with caution. @fig-normalising illustrates the normal density not being directly proportional to its probability of occurrence.
This is concerning, especially at the extreme values, since a small difference in the tail density can have magnitudes of difference in its probability of occurrence.

```{r fig-normalising}
#| fig-cap: "Scatterplot of normal quantiles against their density values. THree tail density values are highlighted with its probability of occurence labelled. Probability is calculated assuming monthly data: with a density of -2, the probability of occurrence is 1/pnorm(-2)/12 = 4 years. The non-linear relationship between the two quantities suggests normalised indexes need to be interpreted with caution since a slight change in the tail distribution can result in magnitudes of difference in its probability of occurrence."
#| fig-width: 9
library(tidyverse)
all <- tibble(x = seq(-3, 3, 0.01), y = pnorm(x)) 
dt <- tibble(x = c(-3, -2.5, -2), y = pnorm(x)) 
label <- tibble(x = c(-3, -2.5, -2), y = pnorm(x), 
                label = c("once in 62 years", "once in 13 years", "once in 4 years"))

dt %>% 
  ggplot() + 
  geom_point(aes(x= x, y = y)) + 
  geom_line(data = all, aes(x = x, y = y)) + 
  ggrepel::geom_label_repel(
    data = label, aes(x = x, y = y, label =label),
    nudge_y = 0.2, nudge_x = 0.1,
    arrow = arrow(length = unit(0.08, "inches")), min.segment.length = 0) +
  scale_x_continuous(breaks = seq(-3, 3, 1)) + 
  xlab("Density") + 
  ylab("Quantile") + 
  theme_bw() + 
  theme(panel.grid.minor = element_blank())
```

### 2022-11-28 paper/paper.tex
section on distribution fit and tweak in normalising plot

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, breakable, sharp corners, enhanced, frame hidden]}{\end{tcolorbox}}\fi

@@

\[F_{\eta}(x(\mathbf{s}; \mathbf{t})),\] where
\(\eta \in H \subseteq \mathbb{R}^{d_{\eta}}\) is the distribution
parameter and \(d_{\eta}\) is the number of parameter of \(\eta\).

Considerations need to be given on selecting the best distribution to
fit, Various discussion on the fitting methods and goodness of fit.
Potentially, more care about how much difference each combination give
in terms of the index value.

sometimes, to account for the monthly variation in the data, the fitting
is done separately for each month and the noramlising step can be
written as \(\Phi^{-1}[F_{\eta}^m(x(\mathbf{s};t_{j^*}))]\), where
\(j^*\) is all the indexes that satisfy \(j^* \mod 12 = m\) for each
\(m = 0, 1,\cdots, 11\).

\begin{itemize}
\tightlist
\item
  In some cases, the distibution and inverse normal can be multivariate,
  i.e.~use copula.
\end{itemize}

@@

This step maps the univariate series into a different scale, typically
for the ease of comparison. For example, a normal scale, {[}0, 1{]},
{[}0, 100{]} may be favored to report certain indexes. In drought
indexes, i.e.~SPI or SPEI, the quantiles from the fitted distribution is
converted into the normal scale via the normal reverse CDF function:
\(\Phi^{-1}(.)\).

Normalising is usually used in the end of the pipeline and its main
difference from the scaling step is that here the change of scale also
changes the distribution of the variable. While being commonly used,
this step can get criticism from analysts for forcing the data into the
decided scale, which can be either unnecessary or inaccurately
exaggerate or downplay the outliers. Also, the use of normal scale needs
to be interpreted with caution. Figure~\ref{fig-normalising} illustrates
the normal density not being directly proportional to its probability of
occurrence. This is concerning, especially at the extreme values, where
a normal density of -5 and -4 has magnitudes of difference into its
probability of occurrence (1 in 3,000,000 and 1 in 31,000,
respectively).

@@

\caption{\label{fig-normalising}scatterplot of quantile vs.~its density
values.}

>>>>

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, frame hidden, interior hidden, sharp corners, boxrule=0pt, enhanced, breakable]}{\end{tcolorbox}}\fi

@@

Distribution fit can be seen as model fitting in its simplest term. It
can be represented by

\[F_{\eta}(x(\mathbf{s}; \mathbf{t})),\] where
\(\eta \in H \subseteq \mathbb{R}^{d_{\eta}}\) is the distribution
parameter and \(d_{\eta}\) is the number of parameter of \(\eta\). A
distribution fit typically aims at finding the distribution that best
fit the data. Analysts may start from a pool of candidate distributions
with a chosen fitting method and goodness of fit measure. While it is
useful to find the ultimate best distribution to fits the data, from a
probabilistic perspective, the fitting procedure itself has uncertainty
associated with the data fed and parameter chosen. A reasonable
alternative is to understand how much the index values can vary given
different distributions, fitting methods, and goodness of fit tests, and
whether these variations are negligible in a given application.

@@

This step maps the univariate series into a different scale, typically
for the ease of comparison across regions. For example, a normal scale,
{[}0, 1{]}, or {[}0, 100{]} may be favored for reporting certain
indexes. In drought indexes, i.e.~SPI or SPEI, the quantiles from the
fitted distribution is converted into the normal scale via the normal
reverse CDF function: \(\Phi^{-1}(.)\).

Normalising is usually used in the end of the pipeline and its main
difference from the scaling step is that here the change of scale also
changes the distribution of the variable. While being commonly used,
this step can get criticism from analysts for forcing the data into the
decided scale, which can be either unnecessary or inaccurately
exaggerate or downplay the outliers. Also, the use of normal scale needs
to be interpreted with caution. Figure~\ref{fig-normalising} illustrates
the normal density not being directly proportional to its probability of
occurrence. This is concerning, especially at the extreme values, since
a small difference in the tail density can have magnitudes of difference
in its probability of occurrence.

@@

\caption{\label{fig-normalising}Scatterplot of normal quantiles against
their density values. THree tail density values are highlighted with its
probability of occurence labelled. Probability is calculated assuming
monthly data: with a density of -2, the probability of occurrence is
1/pnorm(-2)/12 = 4 years. The non-linear relationship between the two
quantities suggests normalised indexes need to be interpreted with
caution since a slight change in the tail distribution can result in
magnitudes of difference in its probability of occurrence.}
